---
title: "Mars Censere Test Notebook"
author: "Richard Offer et al"
output:
  html_notebook: default
  pdf_document:
    latex_engine: xelatex
fontsize: 11pt
mainfont: Arial
sansfont: Arial
monofont: Courier New

params:
  database_path:
    value:
    input: file
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rmarkdown)
library(tidyverse)
library(DBI)
library(gdata)
library(dbplyr)
library(ggplot2)
library(lubridate)
library(reshape2)
library(quantreg)
library(kableExtra)
database_path <- params$database_path #file.choose()
database_info <- file.info(database_path, extra_cols = FALSE)
database_size <- humanReadable(file.size(database_path), units="auto", standard="Unix", digits=0 )
database_mtime <- file.mtime( database_path )
conn <- DBI::dbConnect(RSQLite::SQLite(), database_path)
```

This is a RStudio Test Notebook that is used to see how the simulation is running prior to commiting.

It requires a database that has been already been generated.

```{r load_summary, include=FALSE}
summary_table <- tbl( conn, "summary")
simulations <- tbl( conn, "simulations")
# use collect() to force execution of SQL so we don't carry forward the pending result
# which doesn't cope well with being inserted into running text
num_simulations <- collect(count(simulations))


# Taken from Cookbook for R
# http://www.cookbook-r.com/Graphs/Multiple_graphs_on_one_page_(ggplot2)/

# Multiple plot function
#
# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
# - cols:   Number of columns in layout
# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
#
# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
# then plot 1 will go in the upper left, 2 will go in the upper right, and
# 3 will go all the way across the bottom.
#
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

### Loaded Database Details 

| Value | Details |
|------:|:--------|
| Path | `r database_path` |
| File Size | `r database_size`Bytes |
| Last Modified | `r database_mtime` |
| # Simulations | `r num_simulations$n` |

```{r, echo=FALSE}
sims <-select( collect(simulations), id, initial_mission_lands, mission_ends, final_soldays, final_population, args )

sims$initial_mission_lands <- as_date( sims$initial_mission_lands )
sims$mission_ends <- as_date( sims$mission_ends )

knitr::kable(sims, col.names=c('ID','DateLanded', 'Completed', '#Sols', 'Pop.', 'Args') ) %>%
  column_spec(6, width = "3in")

```

## Database Overview

```{r population_summary, include=FALSE}
summary_plt <- ggplot(data=summary_table, 
              aes(x=ymd_hms(earth_datetime),
                y=population,
                color=simulation_id)
              ) + 
      geom_point(size=0.3) +
      theme(legend.position="none") +
      labs( x = "Date", y="Population", title="Simulation of Martian Population Growth") +
      scale_color_brewer(palette="Paired")
```




```{r quantile, include=FALSE}

gd <- summary_table %>% 
        group_by(earth_datetime) %>% 
        summarise(avg_pop = mean(population), pop_min=min(population), pop_max=max(population), avg_children=mean(children), avg_deaths=mean(deaths), avg_earth_born=mean(earth_born), avg_mars_born=mean(mars_born) )


range_plt <- ggplot(data=gd, 
              aes(x=ymd_hms(earth_datetime), y=avg_pop )
              ) + 
      geom_line( size=0.2 ) +
      geom_smooth( span=0.2, method = "gam", formula = y ~ s(x, bs = "cs") ) +

        theme(legend.position="none") +
      labs( x = "Date", y="Population", title="Mean & Smoothed Mean of Population Growth Across All Simulations") +
      scale_color_brewer(palette="Paired")



```

```{r population_overview, echo=FALSE}
summary_plt
```

Any uneveness at the target population limit is due to the simulations with the highest population count completing their target earlier, and obviously if the simulation with the highest population is removed, then the average will drop. Hense the short term spike and drops.

```{r quantile_overview, echo=FALSE}
range_plt
```

```{r birth_death_rate, include=FALSE}


births_and_deaths <- ggplot(data=summary_table, 
              aes(x=ymd_hms(earth_datetime)) ) + 
      geom_point( aes( y = children ), color="#009E73",  size=0.2) +
      geom_point( aes( y = deaths), color="#000000", size=0.2) +
      theme(legend.position="none") +
      labs( x = "Date", y="Population", title="Births and Deaths") 


births_and_deaths_by_pop <- ggplot(data=summary_table, 
              aes(x=ymd_hms(earth_datetime)) ) + 
      geom_point( aes( y = children / population ), color="#009E73",  size=0.2) +
      geom_point( aes( y = deaths / population), color="#000000", size=0.2) +
      theme(legend.position="none") +
      labs( x = "Date", y="Fraction", title="... as Fraction of Population") 

avg_births_and_deaths <- ggplot(data=gd, 
              aes(x=ymd_hms(earth_datetime)) ) + 
      geom_smooth( aes( y = avg_pop ), color="#56B4E9", span=0.2, method = "gam", formula = y ~ s(x, bs = "cs") ) +
      geom_smooth( aes( y = avg_children ), color="#009E73", span=0.2, method = "gam", formula = y ~ s(x, bs = "cs")) +
      geom_smooth( aes( y = avg_deaths), color="#000000", span=0.2, method = "gam", formula = y ~ s(x, bs = "cs")) +
      theme(legend.position="none") +
      labs( x = "Date", y="Population", title="Births & Deaths Compared to Overall Population") 


avg_earth_and_mars <- ggplot(data=gd, 
              aes(x=ymd_hms(earth_datetime)) ) + 
      geom_smooth( aes( y = avg_earth_born ), color="#56B4E9", span=0.2, method = "gam", formula = y ~ s(x, bs = "cs") ) +
      geom_smooth( aes( y = avg_mars_born ), color="#009E73", span=0.2, method = "gam", formula = y ~ s(x, bs = "cs")) +
      theme(legend.position="none") +
      labs( x = "Date", y="Population", title="Earthers and Martians") 

```

As a fraction in population, you should still see the impact of new missions landing, the sudden influx of adults depresses the fraction of population that are children.

```{r births_and_death_by_pop, echo=FALSE}

multiplot(births_and_deaths, births_and_deaths_by_pop, avg_births_and_deaths, avg_earth_and_mars, layout=matrix(c(1,2,3,3,4,4), nrow=3, byrow=TRUE) )

```

